---
title: "Chapter 3"
author: "Jeff"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidymodels)
library(tidyverse) #note that this doesn't load everything, just the most commonly used packages
library(AppliedPredictiveModeling)
library(corrplot)
library(moments)
```

## Basic computing tools

apropos can be used to scan loaded packages for functions that relate to a certain thing you are looking for. You can use RSiteSearch("confusion",restrict = "functions") to look online, but this will pretty often return a ton of results, especially for something common.

```{r apropos}
apropos("conf_mat") #adapted from "confusionMatrix" in caret
```

```{r load example data}
data("segmentationOriginal") #data for the cell segmentation example project
segData <- subset(segmentationOriginal,Case =="Train")
cellID <- segData$Cell
class <- segData$Class
case <- segData$Case

#remove using tidy
segData <- segData %>% select(-c(Cell,Class,Case)) #adapted from just removing the raw column numbers in base R
# also removing any variables with status in the name
segData <- segData %>% select(-contains("Status")) #first time finding that DPLY is helpful. Base solution with Grep was less trasparent

```

## Transformation

Here I'm trying to see if I can avoid using the e1071 package, which may be pointless. It just seems like a bad idea to get overly comfortable using packages that are designed for specific classes. Here I'm using the moments package instead to calculate skewness

```{r}
skewness(segData$AngleCh1)
skewValues <- apply(segData,2,skewness)
head(skewValues)
```

