---
title: "many_models"
author: "Jeff"
date: "1/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(modelr)
library(tidyverse)
library(tidymodels)
```

## R4DS ch 25 many models

sounds useful for modeling based EDA, so going to work through the examples here too. Also trying to update to be in line with the tidymodels package. 


```{r data}
library(gapminder)
gapminder
```


```{r pressure, echo=FALSE}
gapminder %>% 
  ggplot(aes(year, lifeExp, group = country)) +
    geom_line(alpha = 1/3)

```
```{r Modeling}
# example of model we want to fit, for one country
nz <- filter(gapminder, country == "New Zealand")
nz %>% 
  ggplot(aes(year, lifeExp)) + 
  geom_line() + 
  ggtitle("Full data = ")

nz_mod <- lm(lifeExp ~ year, data = nz)
nz %>% 
  add_predictions(nz_mod) %>%
  ggplot(aes(year, pred)) + 
  geom_line() + 
  ggtitle("Linear trend + ")

nz %>% 
  add_residuals(nz_mod) %>% 
  ggplot(aes(year, resid)) + 
  geom_hline(yintercept = 0, colour = "white", size = 3) + 
  geom_line() + 
  ggtitle("Remaining pattern")
```

```{r nesting data frames}
#this seems like something it would be useful for me to do...

by_country <- gapminder %>%
  group_by(country,continent) %>%
  nest()

country_model <-  function(df) {
  lm(lifeExp ~year, data = df)
}

#use purrr::map() to apply to every element in the list. Because this list contains dfs that works out great for us
by_country <- by_country %>%
  mutate(
    model = map(data, country_model),
    resids = map2(data,model,add_residuals)
    )
by_country

resids <-  unnest(by_country,resids)
resids

resids %>% 
  ggplot(aes(year, resid)) +
    geom_line(aes(group = country),alpha = 1 / 3) + 
    geom_smooth(se = FALSE) +
    facet_wrap(~continent)
         
```

```{r broom to evaluate models}
glance <- by_country %>% 
  mutate(glance = map(model, broom::glance)) %>% 
  unnest(glance)

glance %>%
  arrange(r.squared)

glance %>%
  ggplot(aes(x=continent,y=r.squared)) +
  geom_jitter()

bad_fit <- filter(glance, r.squared < 0.25)

glance %>% 
  filter(r.squared <.25) %>% 
  unnest(data) %>%
  ggplot(aes(year, lifeExp, colour = country)) +
    geom_line()

```
I really love this nesting concept. Thats exactly what I've been looking for.

## Understanding list-columns better
Although you can make list columns directly, they are a little awkward to work with since most functions work with either atomic vectors or data frames. But there are a lot of workflows that can make use of this.
General workflow structure:
1. create a list column using nest(), summarize() + list(), or mutate() + map
2. create other intermediate list columns using map, map2, or pmap. Generally this is something like applying a function (or model) to every list in the initial list column
3. simplify back down to a data frame or vector
  - for single values, use mutate, or map_lgl, map_int, etc
  - for many values, use unnest. (unnest can also be used even if you only want single values, you just end up with extra columns you might not have wanted, and it will cause the other rows to be repeated as much as necessary to yield all the unique rows)

## Other important things broom does
broom::glance(model) returns a row for each model. Each column gives a model summary: either a measure of model quality, or complexity, or a combination of the two.

broom::tidy(model) returns a row for each coefficient in the model. Each column gives information about the estimate or its variability.

broom::augment(model, data) returns a row for each row in data, adding extra values like residuals, and influence statistics.

